---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Ryassofortran

<!-- badges: start -->
<!-- badges: end -->

The goal of Ryassofortran is to provide a convenient way to call the
Fortran90-release of the soil carbon model YASSO15 in R. 

The Fortran-release is on average two orders of magnitude faster than the
R-release. Using the Fortran-release is superior for model calibration purposes.

NOTE: This package is very early in development and is subject to changes.
Furthermore, the package might be removed and included into a more advanced
calibration package in the following months.

## Installation

You can install the development version of Ryassofortran from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("jpusa/Ryassofortran")
```

On Windows systems the installation requires Rtools to be installed.

## Example

This is a basic example that shows you how to run the Fortran-version of YASSO using the package:

```{r example}
library(Ryassofortran)
# Run the YASSO model with sample parameters and data
soilC <- run_yasso(par = sample_parameters, sdl = sample_data_run)
# Show the results
soilC
```
