---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Ryassofortran

<!-- badges: start -->
[![Build Status](https://travis-ci.com/jpusa/Ryassofortran.svg?branch=master)](https://travis-ci.com/jpusa/Ryassofortran)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/jpusa/Ryassofortran?branch=master&svg=true)](https://ci.appveyor.com/project/jpusa/ryassofortran)
<!-- badges: end -->

The goal of Ryassofortran is to provide a convenient way to call the
Fortran90-release of the soil carbon model YASSO15 in R. 

The Fortran-release of YASSO is up to two orders of magnitude faster than the
R-release. The absolute speed increase is not necessarily significant while
simulating soil carbon, but for model calibration purposes the Fortran-release
is superior.

NOTE: This package is very early in development and is subject to changes.
Furthermore, the package might be removed and included into a more advanced
calibration package in the following months.

## Installation

You can install the development version of Ryassofortran from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("jpusa/Ryassofortran")
```

On Windows systems the installation requires Rtools to be installed.

## Examples

Simulate soil carbon with YASSO:

```{r}
library(Ryassofortran)
```

```{r}
# The initial carbon is given as a vector (A, W, E, N, H)
sample_data_run$init
```

```{r example}
# Run the YASSO model with sample parameters and data
soil_c <- run_yasso(
  par = sample_parameters,
  n_runs = sample_data_run$n_runs,
  time = sample_data_run$time,
  weather = sample_data_run$weather,
  init = sample_data_run$init,
  litter = sample_data_run$litter,
  size = sample_data_run$size,
  leac = sample_data_run$leac
)

# Show the results
round(soil_c, 3)
```

Run YASSO in a way intended for calibration:

```{r}
# The initial carbon is given as a matrix
sample_data_cal$init
```

```{r}
# There is a single leaching value for the entire data set
sample_data_cal$leac
```

```{r}
# Run in calibration mode
soil_c_cal <- calibrate_yasso(
  par = sample_parameters,
  n_runs = sample_data_cal$n_runs,
  time = sample_data_cal$time,
  weather = sample_data_cal$weather,
  init = sample_data_cal$init,
  litter = sample_data_cal$litter,
  size = sample_data_cal$size,
  leac = sample_data_cal$leac
)

# Show the results
round(soil_c_cal, 3)
```

