---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Ryassofortran

<!-- badges: start -->
<!-- badges: end -->

The goal of Ryassofortran is to provide convenient R-functions for calling the Fortran90-release of the soil carbon model YASSO15. The Fortran90-release is highly computationally efficient, which makes it ideal for model calibration purposes.


## Installation

### Requirements

* R-version 3.5.0 or higher.
* On Windows systems, Rtools needs to be installed.

You can install the development version of Ryassofortran from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("YASSOmodel/Ryassofortran")
```

## Usage

Ryassofortran provides two R-functions: `run_yasso()` and `calibrate_yasso()`. These functions call the respective Fortran90-wrappers `runyasso` and `calyasso`, which in turn call the Fortran90-subroutine `mod5c` containing the YASSO15 model code. In other words, the package makes it possible to use simple R-functions to run a very fast implementation of YASSO15. While both R-functions essentially call the same model code, there are a few distinctive differences in how they work.

**It is important to explicitly define the data types for the R-function inputs.** The Fortran90-wrappers expect certain data types (double or integer) for certain variables and the code will crash or silently fail if the types are not cast correctly in R.

The `run_yasso()` function is designed for generic use, such as making predictions with YASSO15. The user provides YASSO15 with driver data and initial carbon in a vector. The model "rolls" the carbon forward one time step at a time using the simulated carbon of the current time step as the initial carbon of the next step.

Do soil carbon predictions with `run_yasso()`:

```{r}
library(Ryassofortran)
```

```{r}
# The initial carbon is given as a vector (A, W, E, N, H)
sample_data_run$init
```

```{r example}
# Run the YASSO model with sample parameters and data
soil_c <- run_yasso(
  par = sample_parameters,
  n_runs = sample_data_run$n_runs,
  time = sample_data_run$time,
  temp = sample_data_run$temp,
  prec = sample_data_run$prec,
  init = sample_data_run$init,
  litter = sample_data_run$litter,
  wsize = sample_data_run$wsize,
  leac = sample_data_run$leac
)

# Show the results
round(soil_c, 3)
```

The `calibrate_yasso()` function is highly specialized and not intended for standard use. It is utilized for model calibration at the Finnish Meteorological Institute: In the database used for calibration, there is a measured initial state corresponding to an observed carbon value at each time step. Consequently, the initial carbon is passed to the function as a matrix and the model uses a pre-determined initial state at each time step. Furthermore, the leaching input is a single value instead of a vector, since every calibration dataset has a characteristic leaching.

During calibrations, run YASSO with `calibrate_yasso()`:

```{r}
# The initial carbon is given as a matrix
sample_data_cal$init
```

```{r}
# There is a single leaching value for the entire data set
sample_data_cal$leac
```

```{r}
# Run YASSO during a calibration
soil_c_cal <- calibrate_yasso(
  par = sample_parameters,
  n_runs = sample_data_cal$n_runs,
  time = sample_data_cal$time,
  temp = sample_data_run$temp,
  prec = sample_data_run$prec,
  init = sample_data_cal$init,
  litter = sample_data_cal$litter,
  wsize = sample_data_cal$wsize,
  leac = sample_data_cal$leac
)

# Show the results
round(soil_c_cal, 3)
```

